QDNAseq segmentation

### 2019-10-29 QDNAseq segmentation.Rmd
...

val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
sheets = readxl::excel_sheets(val.file)[8:13]

@@

qdna.dirs = '~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient'
val.train.dists.by.kb = purrr::map( list.files(qdna.dirs, full.names=T), function(dir) {
  do.call(bind_rows, lapply(list.files(dir, 'raw_dist.tsv', full.names = T, recursive = T), function(f) {
    read_tsv(f,col_types = 'ccdddddddd') %>% mutate(kb = basename(dir))  
  }))
})
names(val.train.dists.by.kb) = basename(list.files(qdna.dirs, full.names=T))

@@

```{r}

plist = lapply(train.resids, function(df) {
  lim = c(0, max(df$varMAD_median))
  if (max(df$varMAD_median) > 0.05) lim = c(min(df$varMAD_median), 0.05)
  ggplot(df, aes(varMAD_median)) + geom_histogram(color='lightblue') + labs(title=unique(df$kb)) + xlim(lim)
})
do.call(grid.arrange, c(plist))

@@

>>>>

train.tiles = purrr::map( list.files(qdna.dirs, full.names=T), function(dir) {
  do.call(bind_rows, lapply(list.files(dir, '5e06_tile', recursive = T, full.names = T), function(f) {
    read_tsv(f, col_types = cols( .default = col_double(), X1 = col_character())) %>% dplyr::rename(sample = 'X1') 
  }))
})
names(train.tiles) = basename(list.files(qdna.dirs, full.names=T))

val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
sheets = readxl::excel_sheets(val.file)[8:14]

@@

qdna.dirs = '~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient'
val.dists.by.kb = purrr::map( list.files(qdna.dirs, full.names=T), function(dir) {
  do.call(bind_rows, lapply(list.files(dir, 'raw_dist.tsv', full.names = T, recursive = T), function(f) {
    read_tsv(f,col_types = 'ccdddddddd') %>% mutate(kb = basename(dir))  
  }))
})
names(val.dists.by.kb) = basename(list.files(qdna.dirs, full.names=T))

val.tiles = purrr::map( list.files(qdna.dirs, full.names=T), function(dir) {
  do.call(bind_rows, lapply(list.files(dir, '5e06_tile', recursive = T, full.names = T), function(f) {
    read_tsv(f, col_types = cols( .default = col_double(), X1 = col_character())) %>% dplyr::rename(sample = 'X1') 
  }))
})
names(val.tiles) = basename(list.files(qdna.dirs, full.names=T))

@@

```{r}
plist = lapply(train.resids, function(df) {
  lim = c(0, max(df$varMAD_median))
  if (max(df$varMAD_median) > 0.05) lim = c(min(df$varMAD_median), 0.05)
  ggplot(df, aes(varMAD_median)) + geom_histogram(color='lightblue') + labs(title=unique(df$kb)) + xlim(lim)
})
do.call(grid.arrange, c(plist))

@@

## Post segmentation 5mb bins

```{r}
train.tile.var = do.call(bind_rows, lapply(train.tiles, function(df) {
  df %>% dplyr::summarise_if(is.double, list( ~ sd(.,na.rm=T) ))  
})) %>% add_column(kb  = names(train.tiles), cohort = 'training', .before=1)

val.tile.var = do.call(bind_rows, lapply(val.tiles, function(df) {
  df %>% dplyr::summarise_if(is.double, list( ~ sd(.,na.rm=T) ))  
})) %>% add_column(kb  = names(val.tiles), cohort = 'validation', .before=1)

m = melt(bind_rows(train.tile.var, val.tile.var), id.vars = c('kb','cohort'))

ggplot(m, aes(variable, value, color=cohort, group=cohort)) + ylim(0.95,1.05) + facet_grid(~kb) +  geom_jitter(position = position_dodge(width=1)) + geom_violin(position = position_dodge(width=1), alpha=0.5)

### Post z-adjustment

```{r}
z.sds = lapply(train.tiles, function(df) {
  df %>% dplyr::summarise_if(is.double, list( ~ sd(.,na.rm=T) ))  
})

z.means = lapply(train.tiles, function(df) {
  df %>% dplyr::summarise_if(is.double, list( ~ mean(.,na.rm=T) ))  
})

scaled.training = lapply(train.tiles, function(df) {
  df %>% dplyr::mutate_if(is.double, list(~ BarrettsProgressionRisk:::unit.var(.) ) )
})

scaled.val = lapply(names(val.tiles), function(kb) {
  df = val.tiles[[kb]]
  zm = z.means[[kb]]
  zd = z.sds[[kb]]
  for (col in colnames((df %>% dplyr::select(-sample))))
       df[[col]] = BarrettsProgressionRisk:::unit.var( df[[col]], zm[[col]], zd[[col]]  )
  return(df)       
})
names(scaled.val) = names(val.tiles)

scale.train.tile.var = do.call(bind_rows, lapply(scaled.training, function(df) {
  df %>% dplyr::summarise_if(is.double, list( ~ median(.,na.rm=T) ))  
})) %>% add_column(kb  = names(scaled.training), cohort = 'training', .before=1)

scale.val.tile.var = do.call(bind_rows, lapply(scaled.val, function(df) {
  df %>% dplyr::summarise_if(is.double, list( ~ median(.,na.rm=T) ))  
})) %>% add_column(kb  = names(scaled.val), cohort = 'validation', .before=1)

m = melt(bind_rows(scale.train.tile.var, scale.val.tile.var), id.vars = c('kb','cohort'))

ggplot(m, aes(cohort, value, color=cohort, group=cohort)) + facet_grid(~kb) +  
  geom_jitter(position = position_dodge(width=1)) + 
  geom_violin(position = position_dodge(width=1), alpha=0.5)
